---
title: "p8105_hw5_lq2250"
author: "Lanlan_Qing"
date: "2025-11-10"
output: github_document
---

## Initial Settings
```{r}
library(tidyverse)
library(rvest)
library(plotly)
library(broom)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

**Uniform Distribution Function**

```{r}
# Make the function for returning logistic values for P(X>=2)
birth_uniform = function(n) {
  sample = sample(365, n, replace = TRUE) # randomize birthday samples from 365 days
  n_obs_max = max(table(sample)) # find the mode (represents the highest repeated number)
  n_obs_max >= 2 # return TRUE if appear more than twice
}
```

**Make the data frame and plot**
```{r}
# Make the data frame of group size 2:50 that iterate 10000 times each
birth_df = replicate(10000, map_lgl(2:50, birth_uniform)) |>
  as.data.frame() |>  # convert matrix to data frame
  mutate(group_size = as.factor(2:50)) |>  # add indicate column for size
  relocate(group_size) |>
  pivot_longer(
    "V1":"V10000",
    names_to = "iteration",
    values_to = "judgement"
  ) |>
  select(-iteration) |>
  group_by(group_size, judgement) |> 
  summarise(n_judge = n()) |> # count #TRUE and #FALSE for each group size
  mutate(prob = as.numeric(n_judge/10000), # transfer #observation to prob
         group_size = fct_inorder(str_c("group size = ", group_size))) # Rename group_size

  

# Make a plot showing prob ~ group size
birth_plot = birth_df |>
  plot_ly(x = ~group_size, y= ~prob, color = ~judgement, type = "bar")
```
**Comment**
This is a good simulation result. When group size turns to be small(<23), it is less likely that there are more than two people share the same birthday(FALSE more). As group size increasing (>=23), the probability of people's birthday dates tend to be the same becomes larger(TRUE more). This totally makes sense since there are 365 days. The possibility of given birth at the same day should increases with larger group size.

## Problem 2

1. **Make Normal Function**
```{r}
t_test_func = function(miu){
  sample = rnorm(30, mean = miu, sd = 5) # Generate samples(n=30) from normal distribution N(0,25)
  t.test(sample, alternative = "two.sided", mu = 0, paired = FALSE, conf.level = 0.95) 
  # T-test of each group of samples by two.sided test
}
```

2. **Clean the results and keep necessities**
```{r}
# Generate 5000 times for each miu
t_test_results = map_dfr(1:6, function(miu) {  
  replicate(5000, t_test_func(miu)) |>  # each miu iterate for 5000 times
    map_dfr(broom::tidy) |>  # transfer to data frame
    mutate(miu = miu)  
}) |>  # add column indicating the miu
  relocate(miu) |> 
  select(miu, estimate, p.value) |>

# Reject null if p<0.05
  mutate(reject_null = p.value < 0.05) |>
  select(-p.value)

```

**Clean data for first plot (miu ~ power)**
```{r}
power_df = t_test_results |>
  group_by(miu) |>
# Calculate effect size by miu group
  summarize(mean = mean(estimate),
            s = sd(estimate),
            effect_size = abs(mean)/s,
            power = mean(reject_null),
            reject_null,
            miu) |>
  ungroup() |>
  select(-reject_null) |>
  distinct()
```

**Make Plot!**
```{r}
power_plot <- ggplot(power_df, aes(x = miu, y = power)) +
  geom_point(size = 3) +
  geom_line()
```
**Comment**\
The larger the true mean (effect size), the closer the test power is to 1(the easier it is to reject the null hypothesis).

3. **
```{r}
# Data for all miu_hat
est_df1 = t_test_results |>
  group_by(miu) |>
  summarize(mean = mean(estimate),
            group = "All")

# Data for rejected miu_hat
est_df2 = t_test_results |>
  filter(reject_null == TRUE) |>
  group_by(miu) |>
  summarize(mean = mean(estimate),
            group = "Reject")

# Combine two datasets
mean_est_df = rbind(est_df1, est_df2)

# Make plot
mean_est_plot = ggplot(mean_est_df, aes(x = miu, y = mean, color = group)) +
  geom_point(size = 3) +
  geom_line()+
  geom_abline(intercept = 0, slope = 1, linetype = "dashed")+
  labs(
    x = "True Mean",
    y = "Mean of Estimate Sample Mean"
  )
```
**Comments**
- For all samples, the average estimated mean $\hat{\mu}$ is approximately equal to the true $\mu$ (almost overlap with the reference line y=x), which demonstrates the unbiasedness of the sample mean.

- For samples that reject the null hypothesis, $\hat{\mu}$ is also approximately equal to the true μ. Because the inference of the t-test is "whether there is sufficient evidence to reject H0", and the unbiasedness of the estimator does not rely on the results of the hypothesis test. Even if only the samples that reject H0 are considered, the average estimate can still accurately reflect the true parameters.



